Learning words through recursive pragmatic reasoning about other agents

### 2013-05-30 praglearn.tex
text

@@

@@

\newcommand{\word}{\text{word}}
\newcommand{\obj}{\text{object}}
\begin{align*}
P_{L_0}(\obj | \word) &= \text{lexicon}(\word, \obj) P_{\text{prior}}(\obj | \text{context}). \\
P_{S_i}(\word | \obj) &\propto \exp\Big(\lambda \big(\log P_{L_{i - 1}}(\obj | \word) - \text{cost}(\word)\big)\Big) \\
P_{L_i}(\obj | \word) &\propto P_{S_{i-1}}(\word | \obj) P_{\text{prior}}(\obj | \text{context})
\end{align*}

For the rest of this paper we somewhat arbitrarily fix the softmax parameter to $\lambda = 3$.

>>>>

\newcommand{\word}{\text{word}}
\newcommand{\obj}{\text{object}}

@@

Consider a simple communicative game: two participants each,
separately, view identical arrays of objects. On the
\textit{Speaker's} screen, one object is highlighted; their goal is to
get the \textit{Listener} to click on this item. To do this, they have
available a fixed, finite set of words; they must pick one. The
Listener then receives this word, and attempts to guess which object
the Speaker meant by it.

This suggest a simple probabilistic model. Given any particular
decision procedure used by the Speaker $S$, the optimal way for the
Listener $L$ to judge the probability of each object is to invert this
decision procedure using Bayes' rule, taking into account the prior
likelihood of each object:
\begin{align*}
P_L(\obj | \word) &\propto P_S(\word | \obj) P_{\text{prior}}(\obj |
\text{context})
\end{align*}
$S$, meanwhile, being herself a rational Bayesian agent, chooses her
word so as to soft-maximize a utility function $U(\word | \obj)$. She
gains utility if $L$ is likely to understand her; she loses utility if
the word is particularly costly to say (e.g., it may be long).
\begin{align*}
P_S(\word | \obj) &\propto \exp \lambda U(\word | \obj) \\
U(\word | \obj) &= \log P_L(\obj | \word) - \text{cost}(\word)
\end{align*}
The degree to which she optimizes this utility is controlled by the
softmax parameter $\lambda$. If $\lambda = 1$ and all words cost the
same, then her probability of producing a word will be proportional to
the likelihood that it will cause $L$ to reach the correct
interpretation; as $\lambda$ increases, she becomes more likely to
choose the word which maximizes this likelihood. (We somewhat
arbitrarily set $\lambda = 3$ throughout.)

This model as stated is optimal by various criteria, but suffers from
a fatal flaw: $S$'s decision procedure depends on $L$'s interpretive
procedure, and vice-versa. This means that it can only be implemented
as stated for a very spei

And indeed, in similar games, humans seem to reason only a few levels
deep (XX cite p-beauty contests), and

@@

\begin{align*}
P_{L_0}(\obj | \word) &= \text{lexicon}(\word, \obj) P_{\text{prior}}(\obj | \text{context}). \\
P_{S_i}(\word | \obj) &\propto \exp\Big(\lambda \big(\log P_{L_{i - 1}}(\obj | \word) - \text{cost}(\word)\big)\Big) \\
P_{L_i}(\obj | \word) &\propto P_{S_{i-1}}(\word | \obj) P_{\text{prior}}(\obj | \text{context})
\end{align*}
